### 2015-01-27 comparaison.tex
improvements to section 'Comparison' (Pint + limitations)

\subsection{Evaluation}
To evaluate the effectiveness of our new approach,
we position ourselves with respect to existing methods dealing with different biological models.
We have chosen the following tools, that are detailed below: 
\textsc{GINsim} (Gene Interaction Network Simulation)~\cite{gonzalez2006ginsim,naldi2009logical,naldi2007decision},
\texttt{LIBDDD} (Library of Data Decision Diagrams)~\cite{thierry2009hierarchical,colange2013towards},  PINT \cite{PMR12-MSCS} and the method of Rocca's et al. \cite{roccaasp} which was developed also on ASP.
\annotMF{Il faut trouver un autre nom pour “Rocca's method”.} \repEB{c'est vrai ; "transition state networks with ASP?}
Each method use a specific kind of representation\footnote{The transalation from other models to PH is done through the converters availbale on line: \url{https://github.com/pauleve/pint/tree/master/converters}}:
logical regulatory network for \textsc{GINsim} \footnote{\textsc{GINsim} is a computer tool for the modeling and simulation of genetic regulatory networks \url{http://ginsim.org/} },
Instantiable Transition Systems for \textsc{libDDD} is a library for symbolic model-checking of CTL \& LTL properties. \footnote{\textsc{libDDD} is \url{http://www.lip6.fr/} },
state transition network for Rocca's et al. method
and Process Hitting (PH) for \textsc{PINT} \footnote{\textsc{PINT} is a Process Hitting tool that implements the language, simulation, formal analysis, and translation of PH models. It available online \url{http://loicpauleve.name/pint/} } as well as for our method.

@@

\item \textbf{PINT}
is a library gathering tools and translators related to the PH.
It should be noted that the only reachability analysis developed so far on the Process Hitting
was implemented in the software Pint, and consists in an approximation:
by avoiding to compute the state-transition graph, it is ensured to be really efficient,
at the cost of possibly terminating without being conclusive (although this is rare).
Moreover, in the case of a positive answer,
it currently does not return the execution the path achieving the desired reachability,
but only outputs its conclusion.
In the first example \textsc{PINT} did not give an answer because of the number of goals is greater than 5.
\end{itemize}

\subsection{Limitation}

We developed new methods to check dynamical properties,
namely identifying stable states and finding all possible paths to reach a given goal.
Compared to other methods (Rocca's method, \textsc{GINsim} and \textsc{libddd}),
our method is relatively faster and also permits to study larger networks
(up to $2^{73}$ states in our tests) despite some inconclusive cases.
Indeed, our ASP program is non-conclusive in the cases where the given goal is not
reachable and the model contains loops in its dynamics
(which happens in almost every biological model).
In this case, the program will compute all infinite paths in these loops,
and never reach a goal or a fixed point.
It is still possible, however, to limit the number of iterations to an arbitrary
maximum which will be eventually reached in the case of an endless loop. This is possible with the option \texttt{"--imax=n"}.
\annotMF{On peut éventuellement citer quelques approximations justifiées de ce maximum.}
%%
%\begin{lstlisting}[numbers=none]
%iclingo find\_path.lp <network\_name>.lp -- imax=n
%\end{lstlisting}
%%
Some interesting values can be given to this integer \texttt{n},
which sets the maximum number of steps.
For example, the total number of states is a maximum which is obvious,
as it will never be exceeded by a minimum path,
but too hight to be very interesting.
The total number of sorts can also be an interesting value,
under the hypothesis that each one will change its active process at most once,
which is often the case for Boolean networks,
or, by using the same reasoning, the total number of processes.
In this case, however, a termination with no solution cannot be considered as a formal
negative answer, unless one can prove that the chosen value \texttt{n}
is bigger than the longest possible path in the state-transition graph.
Given our implementation, if the step \texttt{n} is reached,
the computation stops and displays paths that are not valid executions
to the goal.

>>>>

\subsection{Evaluation}
To evaluate the effectiveness of our new approach,
we position ourselves with respect to existing methods dealing with different biological models.
We have chosen the following tools, that are detailed below: 
\textsc{GINsim} (Gene Interaction Network Simulation)~\cite{gonzalez2006ginsim,naldi2009logical,naldi2007decision},
\texttt{LIBDDD} (Library of Data Decision Diagrams)~\cite{thierry2009hierarchical,colange2013towards},  PINT \cite{PMR12-MSCS} and the method of Rocca's et al. \cite{roccaasp} which was developed also on ASP.
\annotMF{Il faut trouver un autre nom pour “Rocca's method”.} \repEB{c'est vrai ; "transition state networks with ASP?}
\annotMF{Je propose plutôt : “ASP for CTL model-checking” qui pourrait se simplifier en “ASP-CTL” dans le tableau.}
Each method use a specific kind of representation\footnote{The transalation from other models to PH is done through the converters availbale on line: \url{https://github.com/pauleve/pint/tree/master/converters}}:
logical regulatory network for \textsc{GINsim} \footnote{\textsc{GINsim} is a computer tool for the modeling and simulation of genetic regulatory networks \url{http://ginsim.org/} },
Instantiable Transition Systems for \textsc{libDDD} is a library for symbolic model-checking of CTL \& LTL properties. \footnote{\textsc{libDDD} is \url{http://www.lip6.fr/} },
state transition network for Rocca's et al. method
and Process Hitting (PH) for \textsc{PINT} \footnote{\textsc{PINT} is a Process Hitting tool that implements the language, simulation, formal analysis, and translation of PH models. It available online \url{http://loicpauleve.name/pint/} } as well as for our method.

@@

\item \textbf{PINT}
is a library gathering tools and translators related to the PH.
It should be noted that Pint contains the only reachability analysis
developed so far natively for the Process Hitting,
before the method proposed in this paper.
It consists in an approximation that avoids to compute the state-transition graph;
It is thus ensured to be really efficient, which explains the fastest results,
but at the cost of possibly terminating without being conclusive (although this is rare).
However, it is not designed for goals consisting of a full state, which
may trigger an exponential process to reach an answer,
which explains the high computing times for some of them.
\todo{Ref vers le tableau}
Moreover, in the case of a positive answer,
it currently does not return the execution the path achieving the desired reachability,
but only outputs its conclusion.
\end{itemize}

\subsection{Strengths and limitations of our method}

In the previous sections,
we developed new methods to check dynamical properties,
namely identifying stable states and finding all possible paths to reach a given goal.
Compared to other methods (Rocca's method, \textsc{GINsim} and \textsc{libddd}),
our method is relatively faster and also permits to study larger networks
(up to $2^{73}$ states in our tests) despite some inconclusive cases.
Indeed, our ASP program is non-conclusive in the cases where the given goal is not
reachable and the model contains loops in its dynamics
(which happens in almost every biological model).
In this case, the program will compute all infinite paths in these loops,
and never reach a goal or a fixed point.
It is still possible, however, to limit the number of iterations to an arbitrary
maximum which will be eventually reached in the case of an endless loop.
This is possible with the option \texttt{"-{}-imax=n"} of \textsc{iclingo},
where \texttt{n} is the maximum number of steps.
%%
%\begin{lstlisting}[numbers=none]
%iclingo find\_path.lp <network\_name>.lp -- imax=n
%\end{lstlisting}
%%
For example, the total number of states is an obvious maximum,
as it will never be exceeded by a minimum path,
but it is too hight to be very interesting.
The total number of sorts is a more interesting value,
under the hypothesis that each one will change its active process at most once,
which is often the case for Boolean networks;
or, with a similar reasoning, the total number of processes can be chosen.
In this case, however, a termination with no solution cannot be considered as a formal
negative answer, unless one can prove that the chosen value \texttt{n}
is bigger than the longest possible path in the state-transition graph.
Given our implementation, if the step \texttt{n} is reached,
the computation stops and displays paths that are not valid executions
to the goal.
\todo{Y a-t-il moyen de les différencier de chemins corrects?}
